---
title: "Waiting Periods and Firearm Suicides"
subtitle: "The Effects of Waiting Periods on Firearm Suicides in the U.S."
author: 
   - Hussain Hadah
   - Gael Compta
# institute: |
#   | Murphy Institute
#   | School of Public Health
date: "12/04/2024"
header-includes:
   - \usepackage{bbm}
   - \usepackage{amsmath}
   - \usepackage{booktabs}
   - \usepackage{siunitx}
   - \newcolumntype{d}{S[input-symbols = ()]}
   - \definecolor{blue}{RGB}{0,114,178}
   - \definecolor{red}{RGB}{213,94,0}
   - \definecolor{yellow}{RGB}{240,228,66}
   - \definecolor{green}{RGB}{0,158,115}
output:
  beamer_presentation:
    slide_level: 3
    toc: false
    theme: "CambridgeUS"
    colortheme: "beaver"
    fonttheme: "structurebold"
    fig_width: 7
    fig_height: 5
    fig_caption: false
    #incremental: true
# bibliography: AttitudesPaper.bib
# biblio-style: apalike
# nocite: '@*'
---

```{r, include=FALSE}
options(tinytex.verbose = TRUE)
library(knitr)
knitr::opts_chunk$set(echo = FALSE)
```

```{r, include=FALSE}
options(tinytex.verbose = TRUE)
library(knitr)
knitr::opts_chunk$set(echo = FALSE)

## Load and install the packages that we'll be using today
if (!require("pacman")) install.packages("pacman")
pacman::p_load(tictoc, parallel, pbapply, future, future.apply, furrr, RhpcBLASctl, memoise, here, foreign, mfx, tidyverse, hrbrthemes, estimatr, ivreg, fixest, sandwich, lmtest, margins, vtable, broom, modelsummary, stargazer, fastDummies, recipes, dummy, gplots, haven, huxtable, kableExtra, gmodels, survey, gtsummary, data.table, tidyfast, dtplyr, microbenchmark, ggpubr, tibble, viridis, wesanderson, censReg, rstatix, srvyr, formatR, sysfonts, showtextdb, showtext, thematic, sampleSelection, RefManageR, DT, googleVis, png, gridExtra, grid, lattice, renderthis, chromote)
# devtools::install_github("thomasp85/patchwork")
remotes::install_github("mitchelloharawild/icons")
remotes::install_github("ROpenSci/bibtex")
remotes::install_github("jhelvy/renderthis", dependencies = TRUE)
remotes::install_github("rstudio/chromote")

# devtools::install_github("ajdamico/lodown")
## My preferred ggplot2 plotting theme (optional)
## https://github.com/hrbrmstr/hrbrthemes
# scale_fill_ipsum()
# scale_color_ipsum()
font_add_google("Fira Sans", "firasans")
font_add_google("Fira Code", "firasans")

showtext_auto()

theme_customs <- theme(
  text = element_text(family = 'firasans', size = 16),
  plot.title.position = 'plot',
  plot.title = element_text(
    face = 'bold', 
    colour = thematic::okabe_ito(8)[6],
    margin = margin(t = 2, r = 0, b = 7, l = 0, unit = "mm")
  ),
)

colors <-  thematic::okabe_ito(5)

# theme_set(theme_minimal() + theme_customs)
theme_set(hrbrthemes::theme_ipsum() + theme_customs)
## Set master directory where all sub-directories are located
mdir <- "/Users/hhadah/Library/CloudStorage/Dropbox/Research/My Research Data and Ideas/SuicideGun"

## Set working directories
pres_trends <- paste0(mdir,"/output/trends")
pres_plots <- paste0(mdir,"/output/figures")
presentation_wd <- paste0(mdir,"/presentation/dec-4")

## Set working directory

# COLOR PALLETES
library(paletteer) 
# paletteer_d("awtools::a_palette")
# paletteer_d("suffrager::CarolMan")

### COLOR BLIND PALLETES
#paletteer_d("colorblindr::OkabeIto")
#paletteer_d("colorblindr::OkabeIto_black")
# paletteer_d("colorBlindness::paletteMartin")
# paletteer_d("colorBlindness::Blue2DarkRed18Steps")
# paletteer_d("colorBlindness::SteppedSequential5Steps")
# paletteer_d("colorBlindness::PairedColor12Steps")
# paletteer_d("colorBlindness::ModifiedSpectralScheme11Steps")
colorBlindness <- paletteer_d("colorBlindness::Blue2Orange12Steps")
cbbPalette <- c("#000000", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")

# scale_colour_paletteer_d("colorBlindness::ModifiedSpectralScheme11Steps", dynamic = FALSE)
# To use for fills, add
  scale_fill_manual(values="colorBlindness::Blue2Orange12Steps")

# To use for line and point colors, add
  scale_colour_manual(values="colorBlindness::Blue2Orange12Steps")
  #<a><button>[Click me](#sources)</button></a>
  
```

# Background

### Stylized facts

- Suicides by firearms constitute the majority of gun deaths in the U.S. (approximately 60\%) 

\bigskip

- Suicides in the U.S. have been on the rise after a steady decline in the 1990s

\bigskip

- Men are more likely to use firearms to commit suicide (more on this in a few slides)

### Suicides in the U.S have been on the rise

```{r plot-1, echo=FALSE, warning=FALSE, out.width="100%", fig.show='hold', fig.align='center'}
library(gridExtra)
# Load your images
p1 <- grid::rasterGrob(readPNG("trends/06-suicide_byyear_popcont.png"))


gridExtra::grid.arrange(
  grobs = list(p1),
  ncol = 1
)

```

### Suicides by firearms have also been on the rise

```{r plot-2, echo=FALSE, warning=FALSE, out.width="100%", fig.show='hold', fig.align='center'}
library(gridExtra)
# Load your images
p1 <- grid::rasterGrob(readPNG("trends/05-suicide_firearm_byyear_popcont.png"))


gridExtra::grid.arrange(
  grobs = list(p1),
  ncol = 1
)

```


### The medical and psychiatric literature

- Previous studies in medicine and psychiatry attempted to study the effect of waiting periods on suicides

- Most of the estimates are non-causal and use cross country analysis

  - e.g. comparing the U.S. to Australia and Europe

### Most major medical and psychiatric organizations support waiting periods to reduce suicides

- The evidence are mixed, and most studies are non-causal

  - Waiting-period requirements were associated with reductions in firearm suicide rates (Edwards et al., 2018; Luca, Malhotra, and Poliquin, 2017)
  - States with Brady Act waiting periods saw firearm suicide declines (Ludwig and Cook, 2000)
  - Gun-related suicides correlated with gun ownership across 21 countries (Killias et al., 2001; Victimization Surveys, N=16–18)
  - Australian gun law reforms reduced firearm suicides but not overall suicides (Baker & McPhedran, 2007; Time-series: 1979–2004)
  - Waiting periods were advocated to mitigate gun-related suicides (Lewiecki & Miller, 2013)
  - Indiana's firearm seizure law was linked to a 7.5% drop in suicides (Kivisto & Phalen, 2018; State-level data: 1981–2015)
  - Gun ownership positively related to firearm suicides in 11 countries (Killias, 1993; Survey, N=28,000)

### Suicide is often considered an impulsive decision that can be prevented by restricting access to means

```{r plot-0, echo=FALSE, warning=FALSE, out.width="80%", fig.show='hold', fig.align='center'}
library(gridExtra)
library(png)  # Make sure this is loaded for readPNG

# Load your images
p1 <- grid::rasterGrob(readPNG("/Users/hhadah/Library/CloudStorage/Dropbox/Research/My Research Data and Ideas/SuicideGun/presentation/dec-4/bottle.png"))
p2 <- grid::rasterGrob(readPNG("/Users/hhadah/Library/CloudStorage/Dropbox/Research/My Research Data and Ideas/SuicideGun/presentation/dec-4/barrier.png"))
p4 <- grid::rasterGrob(readPNG("/Users/hhadah/Library/CloudStorage/Dropbox/Research/My Research Data and Ideas/SuicideGun/presentation/dec-4/golden-gate.png"))
p5 <- grid::rasterGrob(readPNG("/Users/hhadah/Library/CloudStorage/Dropbox/Research/My Research Data and Ideas/SuicideGun/presentation/dec-4/pill-pack.png"))

# Arrange the images in a grid
# You can adjust the layout using ncol and nrow parameters
grid.arrange(p1, p5, p2, p4, ncol=2, nrow=2)
```

### Example from DC

```{r plot-0a, echo=FALSE, warning=FALSE, out.width="80%", fig.align='center'}
library(ggplot2)
library(png)
library(grid)

# Load your image
img <- readPNG("/Users/hhadah/Library/CloudStorage/Dropbox/Research/My Research Data and Ideas/SuicideGun/presentation/dec-4/bridge-barrier.png")

# Create a ggplot with the image
ggplot() +
  annotation_custom(rasterGrob(img), xmin = -Inf, xmax = Inf, ymin = -Inf, ymax = Inf) +
  theme_void() +
  labs(caption = "Source: O'Carroll and Silverman (1994)")
```

### In an Op-Ed, Ronald Reagan supported the passage of the Brady Act and mentioned suicide as a reason

```{r plot-reagan, echo=FALSE, warning=FALSE, out.width="70%", fig.show='hold', fig.align='center'}
library(gridExtra)
library(png)  # Make sure this is loaded for readPNG

# Load your images
p3 <- grid::rasterGrob(readPNG("/Users/hhadah/Library/CloudStorage/Dropbox/Research/My Research Data and Ideas/SuicideGun/presentation/dec-4/reagan.png"))
# Arrange the images in a grid
# You can adjust the layout using ncol and nrow parameters
grid.arrange(p3, ncol=1)
```

### Using the development in the Difference-in-Differences literature, we will answer the following question

- Is there a \textcolor{blue}{causal} effect of \textcolor{red}{waiting periods} on suicides by firearms in the U.S.?

# Data

### Suicide and state-level gun control data

- My primary data source is the National Vital Statistics System (National Center for Health Statistics, 2020)

  - The National Center for Health Statistics collects data for the U.S. government to monitor and improve the nation's death rates 
  - The Multiple Cause of Death files provide the cause of death of every mortality that occurs in the U.S.
  - The data is at the county level and from 1959 to 2019 and has information on sex, age, race, etc.

- For state level gun control policies, I use the RAND state firearm law database (RAND, 2022)

  - This dataset has information on all gun control policies, including waiting periods

\hyperlink{trends-1}{\beamerbutton{Firearm Suicides and Treatment Trends}} \hypertarget{Back2}{}
\hyperlink{trends-2}{\beamerbutton{Suicides and Treatment Trends}} \hypertarget{Back2}{}
\hyperlink{dem-vote}{\beamerbutton{Firearm Suicides by Dem Vote}} \hypertarget{Back2}{}
\hyperlink{density}{\beamerbutton{Firearm Suicides by Density}} \hypertarget{Back2}{}

### Firearms suicide by sex, age, and race

```{r plot-3, echo=FALSE, warning=FALSE, out.width="100%", fig.show='hold', fig.align='center'}
library(gridExtra)
# Load your images
p1 <- grid::rasterGrob(readPNG("trends/221-suicide_firearm_byyear_raceagesex.png"))


gridExtra::grid.arrange(
  grobs = list(p1),
  ncol = 1
)

```

### State-level variation in waiting period policies

```{r plot-4, echo=FALSE, warning=FALSE, out.width="100%", fig.show='hold', fig.align='center'}
library(gridExtra)
# Load your images
p1 <- grid::rasterGrob(readPNG("trends/211-county_year_panel_view.png"))


gridExtra::grid.arrange(
  grobs = list(p1),
  ncol = 1
)
```

### I will use a subsample of states and years where states do not switch back to untreatment

```{r plot-5, echo=FALSE, warning=FALSE, out.width="100%", fig.show='hold', fig.align='center'}
library(gridExtra)
# Load your images
p1 <- grid::rasterGrob(readPNG("trends/192-staggarred_sample_panel_view.png"))


gridExtra::grid.arrange(
  grobs = list(p1),
  ncol = 1
)
```

### Number of clusters per treatment history

```{r plot-6, echo=FALSE, warning=FALSE, out.width="100%", fig.show='hold', fig.align='center'}
library(gridExtra)
# Load your images
p1 <- grid::rasterGrob(readPNG("trends/195-staggarred_sample_panel_view.png"))


gridExtra::grid.arrange(
  grobs = list(p1),
  ncol = 1
)
```

# Empirical Strategy

### Event Study Specification

\begin{align*}
Y_{ist} = & \sum_{l=-K}^{L} \beta_l \mathbbm{1}\{t-E_s=l\} + \theta_i + \lambda_t + \varepsilon_{ist}
\end{align*}

\pause

where:

- $Y_{ist}$ is the number of firearm suicides per 100,000 in county $i$ in state $s$ at time $t$
- $E_s$ is the time period when state $s$ implemented a waiting period
- $\mathbbm{1}\{t-E_s=l\}$ is an indicator variable equal to 1 when time $t$ is $l$ periods away from the implementation of waiting periods in state $s$

### Event Study Specification

\begin{align*}
\textcolor{blue}{Y_{ist}} = & \sum_{l=-K}^{L} \beta_l \textcolor{green}{\mathbbm{1}\{t-\textcolor{red}{E_s}=l\}} + \theta_i + \lambda_t + \varepsilon_{ist}
\end{align*}

where:

- $\textcolor{blue}{Y_{ist}}$ is the number of firearm suicides per 100,000 in county $\textcolor{yellow}{i}$ in state $\textcolor{yellow}{s}$ at time $\textcolor{yellow}{t}$
- $\textcolor{red}{E_s}$ is the time period when state $\textcolor{red}{s}$ implemented a waiting period
- $\textcolor{green}{\mathbbm{1}\{t-\textcolor{red}{E_s}=l\}}$ is an indicator variable equal to 1 when time $t$ is $l$ periods away from the implementation of waiting periods in state $s$


### Event Study Specification (Cont'd)

\begin{align*}
Y_{ist} = & \sum_{l=-K}^{L} \textcolor{blue}{\beta_l} \mathbbm{1}\{t-E_s=l\} + \theta_i + \lambda_t + \varepsilon_{ist}
\end{align*}

\pause

- For $l < 0$, $\textcolor{blue}{\beta_l}$ captures the differences in firearm suicides between treated and control counties before the implementation of waiting periods

\pause

- For $l \geq 0$, $\textcolor{blue}{\beta_l}$ captures the post-treatment effect of waiting periods

- To identify causal estimates of $\textcolor{blue}{\beta_l}$, we need the parallel trends and no anticipation assumptions to hold \hyperlink{attgt}{\beamerbutton{ATT(g, t)}} \hypertarget{Back}{}

### Brady Act and the Parallel Trends Assumption

- The Brady Act was signed into law in 1993 and required a five-day waiting period for the purchase of a handgun

- Supreme Court ruled that the Brady Act was unconstitutional in 1997

```{r plot-brady, echo=FALSE, warning=FALSE, out.width="60%", fig.show='hold', fig.align='center'}
library(gridExtra)
# Load your images
p1 <- grid::rasterGrob(readPNG("trends/218-treated-trends.png"))


gridExtra::grid.arrange(
  grobs = list(p1),
  ncol = 1
)
```

# Results

### I find that waiting periods cause a small reduction in firearm suicides

```{r plot-7, echo=FALSE, warning=FALSE, out.width="100%", out.height="80%" ,fig.show='hold', fig.align='center'}
library(gridExtra)
# Load your images
p1 <- grid::rasterGrob(readPNG("figures/1000-csid-all-fire.png"))


gridExtra::grid.arrange(
  grobs = list(p1),
  ncol = 1
)
```

### I find more significant effects of waiting periods on male firearm suicides

```{r plot-9, echo=FALSE, warning=FALSE, out.width="100%", out.height="80%" ,fig.show='hold', fig.align='center'}
library(gridExtra)
# Load your images
p1 <- grid::rasterGrob(readPNG("figures/1001-csid-men-fire.png"))


gridExtra::grid.arrange(
  grobs = list(p1),
  ncol = 1
)
```

### I find large and significant effects of waiting periods on firearm suicides among individuals aged 55+

```{r plot-10, echo=FALSE, warning=FALSE, out.width="100%", out.height="80%" ,fig.show='hold', fig.align='center'}
library(gridExtra)
# Load your images
p1 <- grid::rasterGrob(readPNG("figures/1002-csid-age_55p-fire.png"))

gridExtra::grid.arrange(
  grobs = list(p1),
  ncol = 1
)
```

### I find large and significant effects of waiting periods on firearm suicides among white individuals

```{r plot-11, echo=FALSE, warning=FALSE, out.width="100%", out.height="80%" ,fig.show='hold', fig.align='center'}
library(gridExtra)
# Load your images
p1 <- grid::rasterGrob(readPNG("figures/1003-csid-white-fire.png"))

gridExtra::grid.arrange(
  grobs = list(p1),
  ncol = 1
)
```

# Results: Non-Firearm Suicides

### I find that waiting periods do increase non-firearm suicides

```{r plot-12, echo=FALSE, warning=FALSE, out.width="100%", out.height="80%" ,fig.show='hold', fig.align='center'}
library(gridExtra)
# Load your images
p1 <- grid::rasterGrob(readPNG("figures/1004-csid-all-other.png"))

gridExtra::grid.arrange(
  grobs = list(p1),
  ncol = 1
)
```

### I find that waiting periods do not increase non-firearm suicides among men

```{r plot-13, echo=FALSE, warning=FALSE, out.width="100%", out.height="80%" ,fig.show='hold', fig.align='center'}
library(gridExtra)
# Load your images
p1 <- grid::rasterGrob(readPNG("figures/1005-csid-men-other.png"))

gridExtra::grid.arrange(
  grobs = list(p1),
  ncol = 1
)
```

### I find that waiting periods do not increase non-firearm suicides among individuals aged 55+

```{r plot-14, echo=FALSE, warning=FALSE, out.width="100%", out.height="80%" ,fig.show='hold', fig.align='center'}
library(gridExtra)
# Load your images
p1 <- grid::rasterGrob(readPNG("figures/1006-csid-age_55p-other.png"))

gridExtra::grid.arrange(
  grobs = list(p1),
  ncol = 1
)
```

### I find that waiting periods do not increase non-firearm suicides among white individuals

```{r plot-15, echo=FALSE, warning=FALSE, out.width="100%", out.height="80%" ,fig.show='hold', fig.align='center'}
library(gridExtra)
# Load your images
p1 <- grid::rasterGrob(readPNG("figures/1007-csid-white-other.png"))

gridExtra::grid.arrange(
  grobs = list(p1),
  ncol = 1
)
```

# Conclusion

### Back of the envelope calculations

- Waiting periods decrease firearm suicides by 0.6 deaths per 100,000 people

  - That is an 8\% decrease in firearm suicides
  - That is a reduction of 1,924 firearm suicides per year

- The benefits of waiting periods could range from  \$11.7 to \$37.9 billion per year
  - Value of statistical life = ranges from \$6.4 to \$19.7 million (HHS, 2024)

### Waiting periods are effective in reducing firearm suicides

- Waiting periods reduce firearm suicides by 0.6 deaths per 100,000 people

- These effects are larger among men, individuals aged 55+, and white individuals

- Waiting periods do not increase non-firearm suicides

- The benefits of waiting periods could be as high as \$37.9 billion per year

###

\Huge Thank you!

# Appendix

### Average Treatment Effect Using Callaway and Sant’Anna (2021)

- I estimate the average treatment effect of waiting periods on firearm suicides by using the not-yet-treated counties as a control group

- The average treatment effect is given by:

\begin{align*}
ATT(g, t) = \mathbb{E}[Y_{is, t} - Y_{is, g-1} \mid G_i = g] - \mathbb{E}[Y_{is, t} - Y_{is, g-1} \mid G_i \in \mathcal{G}_{comp}]
\end{align*}

where:

- $\mathbb{E}[Y_{is, t} - Y_{is, g-1} \mid G_i = g]$ is expected change in outcome for cohort $g$ between periods $g-1$ and $t$
- $\mathbb{E}[Y_{is, t} - Y_{is, g-1} \mid G_i \in \mathcal{G}_{comp}]$ is the expected change in outcome for the comparison group that is not-yet-treated at time $t$
- $G_i$ is the cohort of counties that are treated at time $t$
- $\mathcal{G}_{comp}$ is the set of cohorts that are not-yet-treated at time $t$

\hypertarget{attgt}{}\hyperlink{Back}{\beamerbutton{Back}}

### Average Treatment Effect Using Callaway and Sant’Anna (2021) (Cont'd)

\begin{align*}
ATT(g, t) = \mathbb{E}[Y_{is, t} - Y_{is, g-1} \mid G_i = g] - \mathbb{E}[Y_{is, t} - Y_{is, g-1} \mid G_i \in \mathcal{G}_{comp}]
\end{align*}

where:

- $\mathbb{E}[Y_{is, t} - Y_{is, g-1} \mid G_i = g]$ is expected change in outcome for cohort $g$ between periods $g-1$ and $t$
- $\mathbb{E}[Y_{is, t} - Y_{is, g-1} \mid G_i \in \mathcal{G}_{comp}]$ is the expected change in outcome for the comparison group that is not-yet-treated at time $t$
- $G_i$ is the cohort of counties that are treated at time $t$
- $\mathcal{G}_{comp}$ is the set of cohorts that are not-yet-treated at time $t$
- $\hat{ATT(g, t)} =  \hat{\beta_l}$ if the parallel trends and no anticipation assumptions hold

\hyperlink{Back}{\beamerbutton{Back}}

### Average Treatment Effect Using Callaway and Sant’Anna (2021): Example

\begin{align*}
ATT(g, t) = \mathbb{E}[Y_{is, t} - Y_{is, g-1} \mid G_i = g] - \mathbb{E}[Y_{is, t} - Y_{is, g-1} \mid G_i \in \mathcal{G}_{comp}]
\end{align*}

where:

- $G_i=$ \{Group A (treated in 2020), Group B (treated in 2022), Group C (never treated)\}
- $ATT(A, 2021) = \mathbb{E}[Y_{is, 2021} - Y_{is, 2020} \mid G_i = A] - \mathbb{E}[Y_{is, 2021} - Y_{is, 2020} \mid G_i = B]$
- $ATT(B, 2023) = \mathbb{E}[Y_{is, 2023} - Y_{is, 2022} \mid G_i = B] - \mathbb{E}[Y_{is, 2023} - Y_{is, 2022} \mid G_i = C]$

\hyperlink{Back}{\beamerbutton{Back}}


### Trends in Suicide by Firearms: Treatment and Control States Over Time

```{r plot-app1, echo=FALSE, warning=FALSE, out.width="80%", fig.show='hold', fig.align='center'}
library(gridExtra)
library(png)  # Make sure this is loaded for readPNG

# Load your images
p1 <- grid::rasterGrob(readPNG("trends/01-suicide_firearm_byyearbystate.png"))
# Arrange the images in a grid
# You can adjust the layout using ncol and nrow parameters
grid.arrange(p1, ncol=1)
```


\hypertarget{trends-1}{}\hyperlink{Back2}{\beamerbutton{Back}}

### Trends in Suicide by Firearms: Treatment and Control States Over Time

```{r plot-app2, echo=FALSE, warning=FALSE, out.width="80%", fig.show='hold', fig.align='center'}
library(gridExtra)
library(png)  # Make sure this is loaded for readPNG

# Load your images
p1 <- grid::rasterGrob(readPNG("trends/01-suicide_firearm_byyearbystate.png"))
# Arrange the images in a grid
# You can adjust the layout using ncol and nrow parameters
grid.arrange(p1, ncol=1)
```


\hypertarget{trends-2}{}\hyperlink{Back2}{\beamerbutton{Back}}

### Trends in Suicide by Firearms: By Democratic Vote Share Over Time

```{r plot-app3, echo=FALSE, warning=FALSE, out.width="80%", fig.show='hold', fig.align='center'}
library(gridExtra)
library(png)  # Make sure this is loaded for readPNG

# Load your images
p1 <- grid::rasterGrob(readPNG("trends/219-by-dem-vote.png"))
# Arrange the images in a grid
# You can adjust the layout using ncol and nrow parameters
grid.arrange(p1, ncol=1)
```


\hypertarget{dem-vote}{}\hyperlink{Back2}{\beamerbutton{Back}}

### Trends in Suicide by Firearms: By Density Over Time

```{r plot-app4, echo=FALSE, warning=FALSE, out.width="80%", fig.show='hold', fig.align='center'}
library(gridExtra)
library(png)  # Make sure this is loaded for readPNG

# Load your images
p1 <- grid::rasterGrob(readPNG("trends/220-by-density.png"))
# Arrange the images in a grid
# You can adjust the layout using ncol and nrow parameters
grid.arrange(p1, ncol=1)
```


\hypertarget{density}{}\hyperlink{Back2}{\beamerbutton{Back}}

